### 2008-09-25 doc/client/client_technical.tex
Documentation update from review comments

\begin{enumerate}
\item{The {\TargetGenerator} takes three arguments as input. The first
  argument is a reference to a {\UserConfig} object containing a
  representing the contents of the user's configuration file. The
  second and third arguments contain lists of strings. The first list
  contains individually selected and rejected computing services,
  while the second list contains individually selected and rejected
  index servers. Rejected servers and services are identified by that
  its name is prefixed by a minus sign in the lists. The name of the
  servers and services should be given either in the form of an alias
  defined in the {\UserConfig} object or as the name of its grid
  flavour followed by a colon and the URL of its information contact
  endpoint.}
\item{These lists are parsed through alias resolution before being
  used to initialize the complete list of selected and rejected
  {\URL}s pointing to computing services and index servers.}
\item{An external call is received calling for targets to be
  prepared.}
\item{For each selected index server and computing service a
  {\TargetRetriever} plugin for the server's or service's grid flavour
  is loaded using the ARC loader. The {\TargetRetriever} is
  initialized with its {\URL} and the information about whether it
  represents a computing service or an index server. The call for
  targets is processed by each {\TargetRetriever} in parallel.}
\item{A {\TargetRetriever} representing an index server first tries to
  register at the index server store kept by the {\TargetGenerator}.
  If allowed to register, the index server is queried and the query
  result processed. The {\TargetGenerator} will not allow
  registrations from index servers present in its list of rejected
  index servers or from servers that have already registered
  once. Index servers often registers at more than one index server,
  thus different {\TargetRetriever} may discover the same server.)}
\item{If while processing the query result the {\TargetRetriever}
  finds an other registered index server or a registered computing
  service it creates an a new {\TargetRetriever} for the found server
  or service and forwards the call for targets to the new
  {\TargetRetriever}.}
\item{A {\TargetGenerator} representing a computing service first
  tries to register at the service store kept by the
  {\TargetGenerator}. If allowed to register, the computing server is
  queried and the query result processed. The {\TargetGenerator} will
  not allow registrations from computing services present in its list
  of rejected computing services or from service that have already
  registered once. Computing and services often registers at more than
  one index server, thus different {\TargetRetriever} may discover the
  same service.)}
\item{When processing the query result the {\TargetRetriever} will
  create an {\ExecutionTarget} for each queue found on the computing
  service and collect all possible information about them. It will
  then store the {\ExecutionTarget} in the found targets store kept
  by the {\TargetGenerator} for later usage (e.g. status printing or
  job submission).}
\end{enumerate}

@@

\begin{enumerate}
\item{The {\TargetGenerator} has prepared a list of
  {\ExecutionTarget}s. Depending on the {\URL}s provided to the
  {\TargetGenerator} the list of found {\ExecutionTarget}s may be
  empty or contain several targets. Targets may even represent more
  than one grid flavour. The list of found targets are given as input
  to the {\Broker}.}
\item{In order to rank the found services ({\ExecutionTarget}s) the
  {\Broker} needs detailed knowledge about the job requirements, thus
  the {\JobDescription} is passed as input to the brokering process.}
\item{The {\Broker} outputs a ordered list of {\ExecutionTarget}s
  according to the provided {\JobDescription}.}
\item{Each ExectionTarget has a method to return a specialized
  {\Submitter} which is capable of submitting jobs to the service it
  represents. The best suitable {\ExecutionTarget} for the job is
  asked to return a {\Submitter} for job submission.}
\item{The {\Submitter} takes the {\JobDescription} as input and
  uploads it to the computing service.}
\item{The {\Submitter} identifies local input files from the
  {\JobDescription} and uploads them to the computing service.}
\end{enumerate}

@@

Once a job is submitted job related information (job identification
string, cluster etc) is stored in a local xml file which hosts similar
information for all active jobs. This file may contain jobs running on
completely different grid flavours, and thus job management should be
handled using plugins similar to resource discovery and job
submission. The job managing plugin is called the {\JobController} and
it is supported by the {\JobSupervisor} and {\Job} classes.

@@

\begin{enumerate}
\item{The {\JobSupervisor} takes four arguments as input. The first
  argument is a reference to a {\UserConfig} object containing a
  representing the contents of the user's configuration file. The
  second is a list of strings containing job identifiers and job
  names, the third is a list of strings of clusters to select or
  reject (in the same format as described for the {\TargetGenerator}
  above). The last argument is the name of the file containing the
  local information about active jobs, hereafter called the joblist
  file.}
\item{A job identifier does not uniquely define which grid flavour
  runs a certain job. Thus this information is stored in the joblist
  file upon submission by the {\Submitter} and the joblist file is
  extensively used by the {\JobSupervisor} to identify the
  {\JobController} flavours which are to be loaded. The information in
  the joblist file is also used to lookup the job identifier for jobs
  specified using job names. Alias resolution for the selected and
  rejected clusters are performed using the information in the
  {\UserConfig} object.}
\item{Suitable {\JobController}s are loaded}
\item{The list of job identifiers and selected and rejected clusters
  are passed to each {\JobController} which uses the information to
  fill its internal JobStore.}
\item{Residing within the {\JobSupervisor} the {\JobController}s are
  now accessible for external calls (i.e. job handling).}
\end{enumerate}

@@

The Arc Client Component ({\ACC}) class is a base class needed for the
{\Loader} in order to create loadable classes. It stored information
about which flavour it supports and security related information
needed by all ACC specializations.

@@

 The {\TargetGenerator} class is the main class for resource discovery
 and information retrieval. It loads {\TargetRetriever} plugins in
 accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed}
 (e.g. if an {\URL} pointing to an ARC1 resource is given the
 TargetRetrieverARC1 is loaded).

@@

where \texttt{clusters}, \texttt{indexurls} are lists of strings where
each string is either the name of a grid flavour followed by a colon
and the URL of an information contact endpoint for a computing service
or index server, respectively, of that grid flavour or an alias
defined in the {\UserConfig} object passed as the first
argument, e.g:

@@

where the former is the URL to an index server while the latter is an
URL to a computing service.

@@

The {\TargetRetriever} is base class for {\TargetRetriever} grid
middleware specializations and inherits from the \texttt{ACC} base
class in order to be loadable by the ARC \texttt{Loader}. It is
designed to work in conjunction with the {\TargetGenerator} and
contains the pure virtual method

@@

If an index server query yields an url to a different index server
than the one queried, then the {\TargetRetriever} should call itself
recursively creating a new {\TargetRetriever} for the discovered index
server.

@@

Following a broker decision jobs are to be submitted. Since all
information about the selected computing service resides within the
selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of
returning a {\Submitter} capable of submitting a job to the service
which is represents.

@@

in any supported formats (currently xRSL~\cite{xrsl}, JDL~\cite{jdl},
JSDL~\cite{jsdl}), converts and stores it in a JSDL-like internal job
description format. The identification of the source's description
format is internally handled by a \texttt{JobDescriptionOrderer} class
which identifies the format by pattern matching.

@@

The conversion mainly means a syntactial transformation and the
{\JobDescription} class can of course do nothing in those cases where
not enough data to assemble a given attribute is available or when
information can be lost because the received attribute has no
equivalent in the output language.

@@

No XRSL output generation is implemented yet.

@@

The {\JobController} is a base class for grid specific
specializations, but also the implementor of all public functionality
offered by the {\JobController}s. In other words all virtual functions
of the {\JobController} are private. The initialization of a
(specialized) {\JobController} object takes two steps. First the
{\JobController} specialization for the required grid flavour must be
loaded by the ARC {\Loader}, which sees to that the {\JobController}
receives information about its flavour (grid) and the local joblist
file containing information about all active jobs (flavour
independent). Next step is the filling of the {\JobController}'s job
pool \texttt{JobStore} which is the pool of jobs that the
{\JobController} can manage.

@@

The \texttt{Stat} command prints the job information to screen and in
order to do so the {\JobController} has to query local information
server for the latest status. Due to different protocols used for
different grid flavours (e.g. ldap for production ARC), the
\texttt{GetJobInformation} has to be grid flavour specific and is only
declared as a private virtual method within the {\JobController} base
class. For details about the flavour specific implementations see
Section~\ref{sec:plugins}.

\subsection{{\Job}} The

{\Job} is a generic job class for storing all job related information.
Attributes are derived from the Glue2 information model and thus a
mapping is needed for non Glue2 compliant grid middlewares. Appendix
\ref{app:job} shows the present mapping schema.

@@

\begin{itemize}
\item{If the \texttt{\$X509\_USER\_PROXY} environment variable is set,
  use its value to define a \texttt{ProxyPath} tag in \texttt{cfg}. If
  the file does not exist or can't be read exit with an error.}
\item{Otherwise, if the \texttt{\$X509\_USER\_CERT} environment
  variable is set, use its value and the value of the
  \texttt{\$X509\_USER\_KEY} environment variable to define
  \texttt{CertificatePath} and \texttt{KeyPath} tags in \texttt{cfg}.
  If \texttt{\$X509\_USER\_KEY} is not set or either file can't be
  read exit with an error.}
\item{Otherwise, if the merged user configuration tree (see
  \texttt{ConfTree} containns a \texttt{ProxyPath} tag, copy it to
  \texttt{cfg}.  If the file does not exist or can't be read exit with
  an error.}
\item{Otherwise, if the merged user configuration tree containns a
  \texttt{CertificatePath} tag, copy it to \texttt{cfg} along with the
  accompanying \texttt{KeyPath} tag. If the \texttt{KeyPath} tag is
  undefiend or either filefile does not exist or can't be read exit
  with an error.}
\item{Otherwise, if the file \texttt{/tmp/x509up\_u + userid} exists
  add a \texttt{ProxyPath} tag in \texttt{cfg} pointing to it.}
\item{Otherwise, add \texttt{CertificatePath} and \texttt{KeyPath}
  tags to \texttt{cfg} pointing to
  \texttt{\$HOME/.globus/usercert.pem} and
  \texttt{\$HOME/.globus/userkey.pem}, respectively. If either file
  does not exist exit with error.}
\item{If the \texttt{\$X509\_CERT\_DIR} environment variable is set,
  use its value to define a \texttt{CACertificatesDir} tag in
  \texttt{cfg}. If the directory does not exist, exit with an error.}
\item{Otherwise, if the user ID is not zero and the directory
  \texttt{\$HOME/.globus/certificates} exists add a
  \texttt{CACertificatesDir} tag in \texttt{cfg} pointing to it.}
\item{Otherwise, add a \texttt{CACertificatesDir} tag in \texttt{cfg}
  pointing to \texttt{/etc/grid-security/certificates}. If the
  directory does not exist, exit with an error.}
\end{itemize}

@@

Given all components listed above it possible to write versatile
command line interfaces (cli) for grid job submission and
management. {\arclib} offers the following native commands:

>>>>

\begin{enumerate}
\item{The {\TargetGenerator} takes three arguments as input. The first
  argument is a reference to a {\UserConfig} object containing a
  representation of the contents of the user's configuration file. The
  second and third arguments contain lists of strings. The first list
  contains individually selected and rejected computing services,
  while the second list contains individually selected and rejected
  index servers. Rejected servers and services are identified by that
  its name is prefixed by a minus sign in the lists. The name of the
  servers and services should be given either in the form of an alias
  defined in the {\UserConfig} object or as the name of its grid
  flavour followed by a colon and the URL of its information contact
  endpoint.}
\item{These lists are parsed through alias resolution before being
  used to initialize the complete list of selected and rejected
  {\URL}s pointing to computing services and index servers.}
\item{An external call is received calling for targets to be
  prepared.}
\item{For each selected index server and computing service a
  {\TargetRetriever} plugin for the server's or service's grid flavour
  is loaded using the ARC loader. The {\TargetRetriever} is
  initialized with its {\URL} and the information about whether it
  represents a computing service or an index server. The call for
  targets is processed by each {\TargetRetriever} in parallel.}
\item{A {\TargetRetriever} representing an index server first tries to
  register at the index server store kept by the {\TargetGenerator}.
  If allowed to register, the index server is queried and the query
  result processed. The {\TargetGenerator} will not allow
  registrations from index servers present in its list of rejected
  index servers or from servers that have already registered
  once. Index servers often register at more than one index server,
  thus different {\TargetRetriever}s may discover the same server.}
\item{If while processing the query result the {\TargetRetriever}
  finds an other registered index server or a registered computing
  service it creates a new {\TargetRetriever} for the found server or
  service and forwards the call for targets to the new
  {\TargetRetriever}.}
\item{A {\TargetGenerator} representing a computing service first
  tries to register at the service store kept by the
  {\TargetGenerator}. If allowed to register, the computing server is
  queried and the query result processed. The {\TargetGenerator} will
  not allow registrations from computing services present in its list
  of rejected computing services or from service that have already
  registered once. Computing services often register at more than one
  index server, thus different {\TargetRetriever}s may discover the
  same service.}
\item{When processing the query result the {\TargetRetriever} will
  create an {\ExecutionTarget} for each queue found on the computing
  service and collect all possible information about them. It will
  then store the {\ExecutionTarget} in the found targets store kept
  by the {\TargetGenerator} for later usage (e.g.\ status printing or
  job submission).}
\end{enumerate}

@@

\begin{enumerate}
\item{The {\TargetGenerator} has prepared a list of
  {\ExecutionTarget}s. Depending on the {\URL}s provided to the
  {\TargetGenerator} the list of found {\ExecutionTarget}s may be
  empty or contain several targets. Targets may even represent more
  than one grid flavour. The list of found targets are given as input
  to the {\Broker}.}
\item{In order to rank the found services ({\ExecutionTarget}s) the
  {\Broker} needs detailed knowledge about the job requirements, thus
  the {\JobDescription} is passed as input to the brokering process.}
\item{The {\Broker} outputs a ordered list of {\ExecutionTarget}s
  according to the provided {\JobDescription}.}
\item{Each {\ExecutionTarget} has a method to return a specialized
  {\Submitter} which is capable of submitting jobs to the service it
  represents. The best suitable {\ExecutionTarget} for the job is
  asked to return a {\Submitter} for job submission.}
\item{The {\Submitter} takes the {\JobDescription} as input and
  uploads it to the computing service.}
\item{The {\Submitter} identifies local input files from the
  {\JobDescription} and uploads them to the computing service.}
\end{enumerate}

@@

Once a job is submitted job related information (job identification
string, cluster etc) is stored in a local XML file which hosts similar
information for all active jobs. This file may contain jobs running on
completely different grid flavours, and thus job management should be
handled using plugins similar to resource discovery and job
submission. The job managing plugin is called the {\JobController} and
it is supported by the {\JobSupervisor} and {\Job} classes.

@@

\begin{enumerate}
\item{The {\JobSupervisor} takes four arguments as input. The first
  argument is a reference to a {\UserConfig} object containing a
  representation of the contents of the user's configuration file. The
  second is a list of strings containing job identifiers and job
  names, the third is a list of strings of clusters to select or
  reject (in the same format as described for the {\TargetGenerator}
  above). The last argument is the name of the file containing the
  local information about active jobs, hereafter called the joblist
  file.}
\item{A job identifier does not uniquely define which grid flavour
  runs a certain job. Thus this information is stored in the joblist
  file upon submission by the {\Submitter} and the joblist file is
  extensively used by the {\JobSupervisor} to identify the
  {\JobController} flavours which are to be loaded. The information in
  the joblist file is also used to look up the job identifier for jobs
  specified using job names. Alias resolution for the selected and
  rejected clusters are performed using the information in the
  {\UserConfig} object.}
\item{Suitable {\JobController}s are loaded}
\item{The list of job identifiers and selected and rejected clusters
  are passed to each {\JobController} which uses the information to
  fill its internal \texttt{JobStore}.}
\item{Residing within the {\JobSupervisor} the {\JobController}s are
  now accessible for external calls (i.e. job handling).}
\end{enumerate}

@@

The Arc Client Component ({\ACC}) class is a base class needed for the
{\Loader} in order to create loadable classes. It stores information
about which flavour it supports and security related information
needed by all ACC specializations.

@@

The {\TargetGenerator} class is the main class for resource discovery
and information retrieval. It loads {\TargetRetriever} plugins in
accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed}
(e.g.\ if an {\URL} pointing to an ARC1 resource is given the
TargetRetrieverARC1 is loaded).

@@

where \texttt{clusters} and \texttt{indexurls} are lists of strings
where each string is either the name of a grid flavour followed by a
colon and the URL of an information contact endpoint for a computing
service or index server, respectively, of that grid flavour or an
alias defined in the {\UserConfig} object passed as the first
argument, e.g.:

@@

where the former is the URL to an index server while the latter is an
URL to a computing service. If a string in \texttt{clusters} or
\texttt{indexurls} is prefixed by a minus sign, the corresponding
computing service or index server is rejected and excluded during
resource discovery.

@@

The {\TargetRetriever} is base class for {\TargetRetriever} grid
middleware specializations and inherits from the {\ACC} base class in
order to be loadable by the ARC \texttt{Loader}. It is designed to
work in conjunction with the {\TargetGenerator} and contains the pure
virtual method

@@

If an index server query yields a URL to a different index server
than the one queried, then the {\TargetRetriever} should call itself
recursively creating a new {\TargetRetriever} for the discovered index
server.

@@

Following a broker decision jobs are to be submitted. Since all
information about the selected computing service resides within the
selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of
returning a {\Submitter} capable of submitting a job to the service it
represents.

@@

in any supported formats (currently XRSL~\cite{xrsl}, JDL~\cite{jdl},
JSDL~\cite{jsdl}), converts and stores it in a JSDL-like internal job
description format. The identification of the source's description
format is internally handled by a \texttt{JobDescriptionOrderer} class
which identifies the format by pattern matching.

@@

The conversion mainly means a syntactical transformation and the
{\JobDescription} class can of course do nothing in those cases where
not enough data to assemble a given attribute is available or when
information can be lost because the received attribute has no
equivalent in the output language.

@@

Support for XRSL output generation is currently being developed.

@@

The {\JobController} is a base class for grid specific
specializations, but also the implementer of all public functionality
offered by the {\JobController}s. In other words all virtual functions
of the {\JobController} are private. The initialization of a
(specialized) {\JobController} object takes two steps. First the
{\JobController} specialization for the required grid flavour must be
loaded by the ARC {\Loader}, which sees to that the {\JobController}
receives information about its flavour (grid) and the local joblist
file containing information about all active jobs (flavour
independent). Next step is the filling of the {\JobController}'s job
pool \texttt{JobStore} which is the pool of jobs that the
{\JobController} can manage.

@@

The \texttt{Stat} command prints the job information to screen and in
order to do so the {\JobController} has to query local information
server for the latest status. Due to different protocols used for
different grid flavours (e.g.\ ldap for production ARC), the
\texttt{GetJobInformation} has to be grid flavour specific and is only
declared as a private virtual method within the {\JobController} base
class. For details about the flavour specific implementations see
Section~\ref{sec:plugins}.

\subsection{{\Job}}

The {\Job} is a generic job class for storing all job related
information. Attributes are derived from the Glue2 information model
and thus a mapping is needed for non Glue2 compliant grid middlewares.
Appendix \ref{app:job} shows the present mapping schema.

@@

\begin{itemize}
\item{If the \texttt{\$X509\_USER\_PROXY} environment variable is set,
  use its value to define a \texttt{ProxyPath} tag in \texttt{cfg}. If
  the file does not exist or can't be read exit with an error.}
\item{Otherwise, if the \texttt{\$X509\_USER\_CERT} environment
  variable is set, use its value and the value of the
  \texttt{\$X509\_USER\_KEY} environment variable to define
  \texttt{CertificatePath} and \texttt{KeyPath} tags in \texttt{cfg}.
  If \texttt{\$X509\_USER\_KEY} is not set or either file does not
  exist or can't be read exit with an error.}
\item{Otherwise, if the merged user configuration tree (see
  \texttt{ConfTree} contains a \texttt{ProxyPath} tag, copy it to
  \texttt{cfg}. If the file does not exist or can't be read exit with
  an error.}
\item{Otherwise, if the merged user configuration tree contains a
  \texttt{CertificatePath} tag, copy it to \texttt{cfg} along with the
  accompanying \texttt{KeyPath} tag. If the \texttt{KeyPath} tag is
  undefined or either file does not exist or can't be read exit with
  an error.}
\item{Otherwise, if the file \texttt{/tmp/x509up\_u + userid} exists
  add a \texttt{ProxyPath} tag in \texttt{cfg} pointing to it.}
\item{Otherwise, add \texttt{CertificatePath} and \texttt{KeyPath}
  tags to \texttt{cfg} pointing to
  \texttt{\$HOME/.globus/usercert.pem} and
  \texttt{\$HOME/.globus/userkey.pem}, respectively. If either file
  does not exist exit with error.}
\item{If the \texttt{\$X509\_CERT\_DIR} environment variable is set,
  use its value to define a \texttt{CACertificatesDir} tag in
  \texttt{cfg}. If the directory does not exist, exit with an error.}
\item{Otherwise, if the user ID is not zero and the directory
  \texttt{\$HOME/.globus/certificates} exists add a
  \texttt{CACertificatesDir} tag in \texttt{cfg} pointing to it.}
\item{Otherwise, add a \texttt{CACertificatesDir} tag in \texttt{cfg}
  pointing to \texttt{/etc/grid-security/certificates}. If the
  directory does not exist, exit with an error.}
\end{itemize}

@@

Given all components listed above it is possible to write versatile
command line interfaces (cli) for grid job submission and
management. {\arclib} offers the following native commands:
